\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{A Radial Time Projection Chamber for $\alpha$ detection in CLAS at JLab}

### 2017-05-02 eg6-letters/NIM/EG6_RTPC_NIM.tex
Addition of the paragraph about drift gas and mechanical structure. Plus several minor fixes.

Measurement of the Deeply Virtual Compton Scattering (DVCS) process
($eH \rightarrow e' H' \gamma$, where $H$ is a hadron) necessitates an upgrade
of this setup.  With a 6 GeV electron beam, the majority of DVCS photons are produced
at very forward angles, where the acceptance of the EC is poor. To extend the detection range,
an inner calorimeter (IC) was built for the E01-113 experiment in 2005.
The IC is constructed from 424 lead-tungstate (PbWO$_{4}$) crystals, covering polar 
angles between 5$^{\circ}$ and 15$^{\circ}$. Each crystal is 16 cm long, corresponding
to 17 radiation lengths. The achieved energy resolution is around 3 to 4$\%$ for photon
energies between 2 GeV and 5 GeV, with an angular resolution between 3 and 5 mrad 
\cite{Hyon-suk}. To protect this detector from the large flux of low energy M{\o}ller 
electrons, a 5~T solenoid was placed around the target to shield the IC. 
In addition, the recoil nuclei of the coherent DVCS on helium was detected 
using a RTPC developed to track low energy nuclear fragments. The solenoid field
serving as the analyzing magnet for the particle tracking in the RTPC. This 
setup was used during a three months experimental run
in 2009 with a longitudinally polarized electron beam of {\color{blue} 130 nA 
and energy of 6.064 GeV} incident on a gaseous $^{4}$He target.
{\bf \color{red} Add beam current or luminosity.}

In this paper we report the design, calibration, and performance of the radial 
TPC, organized as follows. Section \ref{sec_design} details the physics 
considerations for our RTPC's design and internal structure. Section 
\ref{sec_readout} describes the properties of the read-out system. Calibration 
strategies are discussed in section \ref{sec_calib}. Finally, the performance 
of the RTPC is described in section \ref{sec_perf}.

@@

We detail here the different regions shown in figure~\ref{fig:RTPC_1_4} starting 
from the beam line and towards larger radius:\\
\begin{itemize}
   \item First, the 6 atm helium-4 target extends along the beam line axis, it 
      is 284~mm long with a 3~mm radius wall made of a 27-$\mu$m-thick Kapton.
   \item The first gas gap covers a radial range from 3 mm to 20 mm. It is 
      filled with $^{4}$He gas at 1~atm to minimize secondary interactions from
      M\o ller electrons scattered by the beam. This 
      region is surrounded by a 4 $\mu$m thick window made of aluminized Mylar 
      and electronically grounded.
   \item The second gap region extends between 20~mm and 30~mm and is filled with a 
      gas mixture of 80$\%$ Neon and 20$\%$ Dimethyl Ether (DME). This region 
      is surrounded by a 4~$\mu$m thick window made of aluminized Mylar, which 
      serves as the cathode and is set at 4260~V.
   \item The drift region is filled with the same Ne-DME gas mixture and extends 
      from the cathode to the first Gas Electron Multiplier(GEM), 60 mm away 
      from the beam axis. The electric field in this region is about 550~V/cm 
      and perpendicular to the beam.
   \item The electron amplification system is composed of three GEMs located at 
      radii of 60, 63 and 66 mm. In this configuration, the first GEM layer 
      serves as the anode and each subsequent GEM is set at a lower voltage to
      obtain a strong ($\sim$1600 V/cm) electric field between the GEM foils. A 
      275~V bias is applied to each GEM for amplification.
   \item The readout board has an internal radius of 69~mm and collects charge
     from the GEMs. Preamplifiers are plugged directly on its outer side and
     transmit the signal to the data acquisition electronics.
\end{itemize}

@@

The GEM technology has been chosen for this RTPC for its flexibility, 
allowing a curved amplification surface. Also, GEMs are known to rarely create 
sparks{\color{blue} \cite{GEM_ref}}, which is important when trying to detect 
highly ionizing slow nuclei that deposit large amount of energy. The GEMs for 
this RTPC are made from a Kapton insulator layer, {\color{blue}50~$\mu$m 
thick}, sandwiched between two 5~$\mu$m copper layers. The mesh of each GEM 
layer is chemically etched with 50~$\mu$m diameter holes with double-conical 
shapes as illustrated in figure~\ref{fig:GEMs}. A potential difference of 400 V 
{\bf \color{red} To be checked and set as above.} is applied between the two 
copper layers of the GEM, creating a very strong electric field in each hole 
leading to high ionization and therefore amplification of the incoming 
electrons. A potential of 150 V {\bf \color{red} To be checked.} is set between 
the GEMs to push the electrons toward the next GEM and eventually the readout 
pads. The gain of each GEM layer is of order 100, making a total gain of about 
$10^{6}$.  {\bf \color{red} We should quote a realistic overall gain or 
nothing.}\\

{\bf \color{red} To be added:

Discuss the choice of drift gas

Discuss the self supporting mechanic of the GEMs.}

@@

We determine the drift speed using tracks reconstructed in the RTPC. In figure 
\ref{fig:RTPC_signals}, a typical $^{4}$He track is represented (in green). After 
it causes ionization in the drift region, the released electrons (in black) 
drift to the cylindrical detection plane under the effect of the electric field. The 
electrons released close to the cathode take the most time to reach the readout 
pads, but cylindrical symmetry insures they always travel the same 
distance. By identifying the maximum TDC measured, we can infer the drift 
speed of the electrons in the RTPC.\\

@@

\begin{figure}[tb]
\hspace*{-1.8cm}
\includegraphics[scale=0.26]{fig/Drift_run_number_1.png}
\caption{$TDC_{Nbmax/2}$ versus the experimental run numbers (time).  } 
\label{fig:Drift_run_number_1}
\end{figure}

@@

\begin{figure}[tb]
\centering
\includegraphics[scale=0.37]{fig/FitResult_p2_10.png}
\caption{$\Delta \phi$ versus TDC distribution in the second pass for tracks
near 35 mm longitudinal position along the RTPC. The black line represents 
the final drift paths in this slice.}
\label{fig:DELTA_PHI_TDC}
\end{figure}

@@

\begin{figure}[tb]
   \centering
\includegraphics[scale=0.350]{fig/EVENT_adc_tdc.png}
\caption{Simulated (upper) and experimental (lower) ADCs and TDCs distributions 
of a track. The colors indicate the pads, same color in top and bottom indicate 
that they are the same pad.}
\label{fig:EVENT_adc_tdc}
\end{figure}

@@

\begin{figure}[tb]\centering
\includegraphics[scale=0.27]{fig/noisy_pad_before_rejection2.png}
\includegraphics[scale=0.27]{fig/noisy_pad_after_rejection2.png}
\caption{The ADC vs. TDC spectrum for an example noisy channel before (top) and after (bottom) noise rejection algorithms.  Only hits associated with tracks are included, and the selection of events and tracks is the same in both plots.}
\label{fig:noise}
\end{figure}

>>>>

Measurement of the Deeply Virtual Compton Scattering (DVCS) process
($eH \rightarrow e' H' \gamma$, where $H$ is a hadron) necessitates an upgrade
of this setup.  With a 6 GeV electron beam, the majority of DVCS photons are produced
at very forward angles, where the acceptance of the EC is poor. To extend the detection range,
an inner calorimeter (IC) was built for the E01-113 experiment in 2005.
The IC is constructed from 424 lead-tungstate (PbWO$_{4}$) crystals, covering polar 
angles between 5$^{\circ}$ and 15$^{\circ}$. Each crystal is 16 cm long, corresponding
to 17 radiation lengths. The achieved energy resolution is around 3 to 4$\%$ for photon
energies between 2 GeV and 5 GeV, with an angular resolution between 3 and 5 mrad 
\cite{Hyon-suk}. To protect this detector from the large flux of low energy M{\o}ller 
electrons, a 5~T solenoid was placed around the target to shield the IC. 
In addition, the recoil nuclei of the coherent DVCS on helium was detected 
using a RTPC developed to track low energy nuclear fragments. The solenoid field
serving as the analyzing magnet for the particle tracking in the RTPC. This 
setup was used during a three months experimental run
in 2009 with a longitudinally polarized electron beam of 130 nA 
and energy of 6.064 GeV incident on a gaseous $^{4}$He target.
{\bf \color{red} Add luminosity.}

In this paper we report the design, calibration, and performance of the radial 
TPC, organized as follows. In section \ref{sec_design}, we detail the design 
and internal structure of the detector. Section \ref{sec_readout} is focused 
on the properties of the read-out system. Calibration 
strategies are discussed in section \ref{sec_calib}. Finally, the performances
of the RTPC are described in section \ref{sec_perf}.

@@

We detail here the different regions shown in figure~\ref{fig:RTPC_1_4} starting 
from the beam line and towards larger radius:\\
\begin{itemize}
   \item First, the 6 atm helium-4 target extends along the beam line axis, it 
      is 284~mm long with a 3~mm radius wall made of a 27-$\mu$m-thick Kapton.
   \item The first gas gap covers a radial range from 3 mm to 20 mm. It is 
      filled with $^{4}$He gas at 1~atm to minimize secondary interactions from
      M\o ller electrons scattered by the beam. This 
      region is surrounded by a 4 $\mu$m thick window made of aluminized Mylar 
      and electronically grounded.
   \item The second gap region extends between 20~mm and 30~mm and is filled with a 
      gas mixture of 80$\%$ neon and 20$\%$ dimethyl ether (DME). This region 
      is surrounded by a 4~$\mu$m thick window made of aluminized Mylar, which 
      serves as the cathode and is set at 4260~V.
   \item The drift region is filled with the same Ne-DME gas mixture and extends 
      from the cathode to the first Gas Electron Multiplier(GEM), 60 mm away 
      from the beam axis. The electric field in this region is about 550~V/cm 
      and perpendicular to the beam.
   \item The electron amplification system is composed of three GEMs located at 
      radii of 60, 63 and 66 mm. In this configuration, the first GEM layer 
      serves as the anode and each subsequent GEM is set at a lower voltage to
      obtain a strong ($\sim$1600 V/cm) electric field between the GEM foils. A 
      275~V bias is applied to each GEM for amplification.
   \item The readout board has an internal radius of 69~mm and collects charge
     from the GEMs. Preamplifiers are plugged directly on its outer side and
     transmit the signal to the data acquisition electronics.
\end{itemize}

@@

The GEM technology has been chosen for this RTPC for its flexibility, 
allowing a curved amplification surface. Also, GEMs are known to rarely create 
sparks{\color{blue} \cite{GEM_ref}}, which is important when trying to detect 
highly ionizing slow nuclei that deposit large amount of energy. The GEMs for 
this RTPC are made from a Kapton insulator layer, {\color{blue}50~$\mu$m 
thick}, sandwiched between two 5~$\mu$m copper layers. The mesh of each GEM 
layer is chemically etched with 50~$\mu$m diameter holes with double-conical 
shapes as illustrated in figure~\ref{fig:GEMs}. A potential difference of 275 V 
is applied between the two 
copper layers of the GEM, creating a very strong electric field in each hole 
leading to high ionization and therefore amplification of the incoming 
electrons. A high potential of about 1600 V/cm is set between 
the GEMs to push the electrons toward the next GEM and eventually the readout 
pads. \\

The drift gas used in the experiment is a 80-20\% neon-DME mixtur, this choice 
has been made in order to balance the good drift properties with a reasonnable
Lorentz angle. Calculations using the MAGBOLTZ program~\cite{MAGBOLTZ} 
showed that with the perpendicular 5~T field, we would have a Lorentz angle of 
about {\bf \color{red} (To be added) } with this mixture. As can be seen, the
Lorentz angle is already significant, a larger angle would increase significantly
the length of the drift paths as the electron would spiral around the beam line.
{\bf \color{red} (To be continued commenting on higher/lower proportion of gas 
and other gas) }

One of the important problematic to develop the radial TPC was to obtain a 
good support structure of the GEM foils, allowing an easy installation of the 
GEM in good conditions and with high precision. At the same time, we wanted to 
keep the material budget small in the forward region where we detect other 
particles in subsequent detectors. We successfuly realized these prerequisites 
by using fiber glass rings glued to each ends of the GEM foils to form self 
supporting cylinders that could be installed independently in the RTPC after 
all gluing and soldering operations.  Because the rigidity of the high 
rigidity of the GEM foil, the strcture was selfsupporting and only the 
backward part of the cylinder was fixed to the main mechanical structure. This 
design only left a light fiberglass ring in the forward region, reducing to a 
minimum secondary interactions.

@@

We determine the drift speed using tracks reconstructed in the RTPC. In figure 
\ref{fig:RTPC_signals}, a typical $^{4}$He track is represented (in green). After 
it causes ionization in the drift region, the released electrons (in black) 
drift to the cylindrical detection plane under the effect of the electric field. The 
electrons released close to the cathode take the most time to reach the readout 
pads, but cylindrical symmetry insures they always travel the same 
distance. By identifying the maximum time T measured, we can infer the drift 
speed of the electrons in the RTPC.\\

@@

\begin{figure}[tb]
\hspace*{-1.8cm}
\includegraphics[scale=0.26]{fig/Drift_run_number_1.png}
\caption{$T_{Nbmax/2}$ versus the experimental run numbers (time).  } 
\label{fig:Drift_run_number_1}
\end{figure}

@@

\begin{figure}[tb]
\centering
\includegraphics[scale=0.37]{fig/FitResult_p2_10.png}
\caption{$\Delta \phi$ versus T distribution in the second pass for tracks
near 35 mm longitudinal position along the RTPC. The black line represents 
the final drift paths in this slice.}
\label{fig:DELTA_PHI_TDC}
\end{figure}

@@

\begin{figure}[tb]
   \centering
\includegraphics[scale=0.350]{fig/EVENT_adc_tdc.png}
\caption{Simulated (upper) and experimental (lower) ADC and T distributions 
of a track. The colors indicate the pads, same color in top and bottom indicate 
that they are the same pad.}
\label{fig:EVENT_adc_tdc}
\end{figure}

@@

\begin{figure}[tb]\centering
\includegraphics[scale=0.27]{fig/noisy_pad_before_rejection2.png}
\includegraphics[scale=0.27]{fig/noisy_pad_after_rejection2.png}
\caption{The ADC vs. T spectrum for an example noisy channel before (top) and after (bottom) noise rejection algorithms.  Only hits associated with tracks are included, and the selection of events and tracks is the same in both plots.}
\label{fig:noise}
\end{figure}
